---
title: "Summarize results of traditional and advanced approaches in a table"
author: "Charlotte Meinke"
date: "9/8/2023"
output: html_document
---

# Load packages
```{r}
library(pander)
library(dplyr)
library(purrr)
library(tibble)
```

# Advanced approach: Import data
## Set paths for advanced approaches
```{r}
# Set paths for folders of advanced approaches
path_base = "Y:\\PsyThera\\Projekte_Meinke\\PAI_Personalized_Advantage_Index\\Results_PAI_Test"
paths_advanced <- list(
  "panic.adv\n(ridge)" = file.path(path_base, "PANIK_net_ridge_regression"),
  "panic.adv\n(ridgewhp)" = file.path(path_base,"PANIK_net_ridge_regression_hp_tuned_grid"),
  "panic.adv\n(rf)" = file.path(path_base, "PANIK_net_random_forest"),
  "protect.adv\n(ridge)" = file.path(path_base,"PROTECT_AD_ridge_regression"),
  "protect.adv\n(ridgewhp)" = file.path(path_base,"PROTECT_AD_ridge_regression_hp_tuned_grid"),
  "protect.adv\n(rf)" = file.path(path_base, "PROTECT_AD_random_forest"),
  "aprotectsympscore.adv\n(ridge)" = file.path(path_base,"PROTECT_AD_symptom_ridge_regression"),
  "aprotectfuncscore.adv\n(ridge)" =  file.path(path_base,"PROTECT_AD_function_ridge_regression")
)
```

## Function: Import results of advanced approaches
```{r}
# import_model_summary_advanced <- function(file_path, model, name_modelperformance_file, name_PAI_file){
#   
#   # Load model performance metrics
#   path_model_eval = file.path(file_path,name_modelperformance_file)
#   df_model_eval = read.delim(path_model_eval)
#   colnames(df_model_eval) = c("Metric",model)
#   
#   # Load PAI evaluation metrics
#   path_PAI_eval = file.path(file_path,name_PAI_file)
#   df_PAI_eval = read.delim(path_PAI_eval, header = TRUE)
#   
#   # Transpose PAI evaluation metrics
#   df_PAI_eval$X <- NULL
#   df_PAI_eval_t <- as.data.frame(t(df_PAI_eval))
#   df_PAI_eval_t$Metric <- rownames(df_PAI_eval_t)
#   colnames(df_PAI_eval_t) <- c(model,"Metric")
#   # Merge by metric
#   df = rbind(df_model_eval,df_PAI_eval_t)
#   return(df)
# }

import_model_summary_advanced <- function(file_path, model, name_modelperformance_file, name_PAI_file){
  
  # Initialize an empty dataframe
  df <- data.frame(Metric = character(), stringsAsFactors = FALSE)
  
  # Check if name_modelperformance_file is provided
  if (!missing(name_modelperformance_file)) {
    # Load modal evaluation metrics
    path_model_eval = file.path(file_path, name_modelperformance_file)
    df_model_eval = read.delim(path_model_eval)
    colnames(df_model_eval) = c("Metric", model)
    
    # Append model evaluation metrics to the dataframe
    df <- rbind(df, df_model_eval)
  }
  
  # Load PAI evaluation metrics
  path_PAI_eval = file.path(file_path, name_PAI_file)
  df_PAI_eval = read.delim(path_PAI_eval, header = TRUE)
  
  # Transpose PAI evaluation metrics
  df_PAI_eval$X <- NULL
  df_PAI_eval_t <- as.data.frame(t(df_PAI_eval))
  df_PAI_eval_t$Metric <- rownames(df_PAI_eval_t)
  colnames(df_PAI_eval_t) <- c(model, "Metric")
  
  # Append PAI evaluation metrics to the dataframe
  df <- rbind(df, df_PAI_eval_t)
  
  return(df)
}
```

## Import and save as dataframe
```{r}

results_files_advanced_all <- list()
for(model in names(paths_advanced)){
  results_files_advanced_all[[model]] = import_model_summary_advanced(paths_advanced[[model]], 
                                                              model,
                                                              name_modelperformance_file = "modelperformance_summary.txt",
                                                              name_PAI_file = "PAI_summary_all.txt")
}

results_files_advanced_50perc <- list()
for(model in names(paths_advanced)){
  results_files_advanced_50perc [[model]] = import_model_summary_advanced(paths_advanced[[model]], 
                                                              model,
                                                              name_PAI_file = "PAI_summary_50_perc.txt")
}

```

## Function: Summarize results of advanced approach
```{r}
process_results_advanced <- function(results_files_advanced, mean_metrics, sd_metrics,add_SD_brackets = TRUE) {
  # This function collects the important results of the advanced approaches in a dataframe
  # mean_metrics = names of mean metrics of interest
  # sd_metrics = names of corresponding sds
  # Reduce results_files_advanced by inner join on "Metric"
  result <- results_files_advanced %>% 
    reduce(inner_join, by = "Metric") %>%
    select(Metric, everything())
  
  # Filter the results based on mean_metrics and sd_metrics
  results_adv_df <- result[result$Metric %in% c(mean_metrics, sd_metrics), ]
  
  # Transpose dataframe
  results_t <- as.data.frame(t(results_adv_df))
  new_header <- as.character(results_t[1, ])  # Extract second row as character vector
  results_t <- results_t[-c(1), ]  # Remove the first and second rows
  colnames(results_t) <- new_header
  results_t[] <- lapply(results_t, as.numeric)
  
  # Add standard deviations in brackets to mean columns
  if (add_SD_brackets == TRUE){
    for (i in seq_along(mean_metrics)) {
      mean <- mean_metrics[i]
      sd <- sd_metrics[i]
      results_t[[mean]] <- paste(round(results_t[[mean]], 2), " (", round(results_t[[sd]], 2), ")", sep = "")
      results_t[[sd]] <- NULL
    }
  } 
  
  # Retranspose
  results_adv_df <- as.data.frame(t(results_t))
  
  # Add metrics as first columns (row names in results_t)
  results_adv_df <- rownames_to_column(results_adv_df, var = "Metric")
  
  return(results_adv_df)
}
```

## Summarize results
```{r }
mean_metrics <- c("Mean_all_correlation","Mean_all_RMSE","Mean_all_MAE","mean.abspai","mean.Cohens.d")
sd_metrics <- c("Std_all_correlation","Std_all_RMSE","Std_all_MAE","mean.abspai.SD", "mean.Cohens.d.SD")
results_adv_df <- process_results_advanced(results_files_advanced = results_files_advanced_all, 
                                           mean_metrics = mean_metrics,
                                           sd_metrics = sd_metrics)

mean_metrics_50perc <- c("mean.abspai","mean.Cohens.d")
sd_metrics_50perc <- c("mean.abspai.SD", "mean.Cohens.d.SD")
results_adv_df_50perc <- process_results_advanced(results_files_advanced = results_files_advanced_50perc, 
                                                  mean_metrics = mean_metrics_50perc,
                                                  sd_metrics = sd_metrics_50perc)
results_adv_df_50perc$Metric <- paste(results_adv_df_50perc$Metric,"50perc",sep = ".")

results_adv_df_50perc
```

# Traditional approach: Import and save as dataframe
## Set paths
```{r}
path_trad <- list(
  "panic.trad\n(linear)"= file.path(path_base,"PANIK_traditional_approach"),
  "protect.trad\n(linear)"= file.path(path_base,"Protect_traditional_approach")
)
```

## Function: Import results of traditional approach
```{r}

# import_model_summary_traditional <- function(file_path, model, name_modelperformance_file, name_PAI_file){
#   
#   # Load model evaluation metrics
#   path_model_eval = file.path(file_path,name_modelperformance_file)
#   df_model_eval = read.delim(path_model_eval)
#   #colnames(df_model_eval) = c("Metric",model)
#   
#   # Load PAI evaluation metrics
#   path_PAI_eval = file.path(file_path,name_PAI_file)
#   df_PAI_eval = read.delim(path_PAI_eval, header = TRUE)
#   
#   # Concat both dfs
#   df = cbind(df_model_eval, df_PAI_eval)
#   
#   # Transpose df
#   df_t <- as.data.frame(t(df))
#   df_t$Metric <- rownames(df_t)
#   colnames(df_t) <- c(model,"Metric")
#   
#   return(df_t)
# }

import_model_summary_traditional <- function(file_path, model, name_modelperformance_file, name_PAI_file){
  
  # Check if name_modelperformance_file is provided
  if (!missing(name_modelperformance_file)) {
    # Load model evaluation metrics
    path_model_eval = file.path(file_path, name_modelperformance_file)
    df_model_eval = read.delim(path_model_eval)
  }
  
  # Load PAI evaluation metrics
  path_PAI_eval = file.path(file_path, name_PAI_file)
  df_PAI_eval = read.delim(path_PAI_eval, header = TRUE)
  
  if (!missing(name_modelperformance_file)) {
    df <- cbind(df_model_eval, df_PAI_eval)
  }else
    df <-  df_PAI_eval
  
  # Transpose df
  df_t <- as.data.frame(t(df))
  df_t$Metric <- rownames(df_t)
  colnames(df_t) <- c(model, "Metric")
  
  return(df_t)
}
```

## Import
```{r}
results_trad <- list()
for(model in names(path_trad)){
  results_trad[[model]] = import_model_summary_traditional(file_path = path_trad[[model]], 
                                                           model = model,
                                                           name_modelperformance_file = "model_eval_summary.txt",
                                                           name_PAI_file = "PAI_eval_summary.txt")
}

results_trad_50perc <- list()
for(model in names(path_trad)){
  results_trad_50perc[[model]] = import_model_summary_traditional(path_trad[[model]], 
                                                           model,
                                                           name_PAI_file = "PAI_eval_50_perc.txt")
}
```

## Summarize results
```{r}
results_trad_df <- results_trad %>%
  reduce(inner_join, by = "Metric")

results_trad_df_50perc <- results_trad_50perc %>%
  reduce(inner_join, by = "Metric")
results_trad_df_50perc$Metric <- paste(results_trad_df_50perc$Metric,"50perc",sep = ".")

results_trad_all <- rbind(results_trad_df,results_trad_df_50perc)
```

# Combine traditional and advanced dataframes
## Recode variable names of advanced results
```{r}
# Names of metrics in the traditional and advanced approach are originally different as the advanced approach calculates the mean across folds/repetitions
# Therefore, metrics need to be renamed in order to merge both dataframes
#results_adv_df$Metric <- as.factor(results_adv_df$Metric)

results_adv_df$Metric <- recode(results_adv_df$Metric, "Mean_all_correlation" = "correlation", "Mean_all_RMSE" = "RMSE", 
                    "Mean_all_MAE" = "MAE", "mean.Cohens.d" = "cohens.d")
results_adv_df_50perc$Metric <- recode(results_adv_df_50perc$Metric, "mean.Cohens.d.50perc" = "cohens.d.50perc")
results_adv_all <- rbind(results_adv_df, results_adv_df_50perc)
```

## Join dataframes
```{r}
df <- merge(results_trad_all, results_adv_all, by = "Metric")
```

# Make dataframe nicer
## Order columns and round 
```{r}
names_advanced <- names(paths_advanced)
names_panic <- names_advanced[grep("^panic", names_advanced)]
names_protect <- names_advanced[grep("^protect.", names_advanced)]
names_protect_alternatives <- names_advanced[grep("score",names_advanced)]

cols_nice_order <- c("Metric",names(path_trad)[1],names_panic,names(path_trad)[2],names_protect,names_protect_alternatives)
df <- df[, c(cols_nice_order)]

df$`panic.trad\n(linear)` <- round(as.numeric(df$`panic.trad\n(linear)`),2)
df$`protect.trad\n(linear)` <- round(as.numeric(df$`protect.trad\n(linear)`),2)
```

## Rename Column and Row names (= Headings)
```{r}
library(stringr)
df_renamed <- df

# Rename datasets in column names
old_substrings <- c("panic", "aprotect", "protect","funcscore","sympscore","whp")
new_substrings <- c("PANIC-net", "Protect-AD", "Protect-AD"," function"," symptom"," w hp")
cols_nice_order_renamed <- str_replace_all(cols_nice_order, setNames(new_substrings, old_substrings))
colnames(df_renamed) <- cols_nice_order_renamed

# Get names of columns
names_panic_new <- cols_nice_order_renamed[grep("^PANIK", cols_nice_order_renamed)]
names_protect_new <- cols_nice_order_renamed[grep("^Protect.", cols_nice_order_renamed)]
names_protect_alternatives_new <- cols_nice_order_renamed[grep("function|symptom",cols_nice_order_renamed)]

# Rename and order metrics
df_renamed$Metric <- factor(df_renamed$Metric, levels = c("cohens.d", "cohens.d.50perc","mean.abspai", "mean.abspai.50perc","MAE","RMSE","correlation"))
levels(df_renamed$Metric) <- c("Cohen´s d","Cohen´s d 50%", "Mean PAI","Mean PAI 50%","MAE","RMSE","Correlation")
df_renamed <- df_renamed[order(df_renamed$Metric), ]
```

# Generate apa-table
```{r}
# Rows: correlation, MAE, RMSE, Mean PAI, Cohen´s optimal vs. nonoptimal
library(flextable)

# Set defaults
set_flextable_defaults(
padding.bottom = 3,
padding.top = 3,
padding.left = 0.5,
paddings.right = 0.5,
theme_fun = "theme_apa",
font.family = "Arial",
font.size = 11
)

ft <- flextable(df_renamed)

# Add space between Panik and PROTECT
ft <- flextable(df_renamed, col_keys = c("Metric",names_panic_new,"blank",names_protect_new))


# ft <- add_header_row(ft, values = c("","traditional approach","revised\napproach","traditional approach", "revised approach"),
#                      colwidth = c(1,1,2,1,1), top = TRUE)
# ft <- add_header_row(ft, values = c("","Panic Net dataset","PROTECT-AD dataset"),
#                      colwidths = c(1,2,3))


# Split headings
ft <- separate_header(x = ft, opts = c("span-top"), split = "[.]")


# Make sure that the first column is left-aligned (APA style!) and make the space empty
ft <- ft %>% align(j = 1,align = "left") %>% empty_blanks()

ft

# Save as ft 
library(officer)

margins <- page_mar(
  bottom = 0.5,
  top = 0.5,
  right = 0.5,
  left = 0.5,
  header = 0.5,
  footer = 0.5,
  gutter = 0.5
)

format_table <- prop_section(
  page_size = page_size(orient = "wide"),
  page_margins = margins)

save_as_docx(
  ft,
  path = "Table_2_ML_performance.docx", pr_section = format_table)

#C:\\Users\\charl\\Documents\\Promotion\\PAI-Paper\\plots\\
```

# Create nice plot for poster
## Summarize results of advanced approach for poster
```{r}
mean_metrics <- c("Mean_all_correlation","Mean_all_RMSE","Mean_all_MAE","mean.abspai","mean.Cohens.d")
sd_metrics <- c("Std_all_correlation","Std_all_RMSE","Std_all_MAE","mean.abspai.SD", "mean.Cohens.d.SD")
results_adv_df_poster <- process_results_advanced(results_files_advanced = results_files_advanced_all, 
                                           mean_metrics = mean_metrics,
                                           sd_metrics = sd_metrics,
                                           add_SD_brackets = FALSE)
```
## Combine traditional and advanced results
### Recode variable names of advanced results
```{r}
# Names of metrics in the traditional and advanced approach are originally different as the advanced approach calculates the mean across folds/repetitions
# Therefore, metrics need to be renamed in order to merge both dataframes
#results_adv_df$Metric <- as.factor(results_adv_df$Metric)

results_adv_df_poster$Metric <- recode(results_adv_df_poster$Metric, "Mean_all_correlation" = "correlation", "Mean_all_RMSE" = "RMSE", 
                    "Mean_all_MAE" = "MAE", "mean.Cohens.d" = "cohens.d")
```
### Add sd_metrics to traditional results 
```{r}
colnames(results_trad_df)
# Initialize a list to store new rows
new_rows <- list()

# Vector of sd_metrics
sd_metrics <- c("Std_all_correlation", "Std_all_RMSE", "Std_all_MAE", "mean.abspai.SD", "mean.Cohens.d.SD")

# Loop through the metrics
for (sd_metric in sd_metrics) {
  # Create a new row with appropriate structure (assuming 3 columns)
  new_row <- c(0, sd_metric, 0)
  # Store the new row in the list
  new_rows[[length(new_rows) + 1]] <- new_row
}

# Combine all the new rows into a matrix or data frame
new_rows_df <- do.call(rbind, new_rows)

new_rows_df <- as.data.frame(new_rows_df, stringsAsFactors = FALSE)
colnames(new_rows_df) <- colnames(results_trad_df)

# Bind the new rows to the existing dataframe
results_trad_df_poster <- rbind(results_trad_df, new_rows_df)
```
### Combine traditional and advanced df
```{r}
df <- merge(results_trad_df_poster, results_adv_df_poster, by = "Metric")
```
## Reduce df
```{r}
metrics_interest <- c("cohens.d","mean.Cohens.d.SD")
df_plot <- df[df$Metric %in% metrics_interest,]
df_plot$Metric <- gsub("cohens.d","Metric_cohens.d",df_plot$Metric)
df_plot$Metric <- gsub("mean.Cohens.d.SD","Metric_mean.Cohens.d.SD",df_plot$Metric)
```

## Turn into long-format
```{r}
approaches <- colnames(df_plot)[colnames(df_plot)!= "Metric"]

df_plot_long_1 <- reshape(data=df_plot[df_plot$Metric == "Metric_cohens.d",], varying = approaches,
          v.name=c("Cohens_d"),
          times= approaches,
          timevar = "approach",
          direction="long")

df_plot_long_1$Metric <- NULL
df_plot_long_1$id <- NULL

df_plot_long_2 <- reshape(data=df_plot[df_plot$Metric == "Metric_mean.Cohens.d.SD",], varying = approaches,
          v.name=c("SD"),
          times= approaches,
          timevar = "approach",
          direction="long")

df_plot_long_2$Metric <- NULL
df_plot_long_2$id <- NULL

df_plot_long <- merge(df_plot_long_1, df_plot_long_2, by = "approach")
```
## Prepare data for plot
```{r}
df_plot_long$Cohens_d <- as.numeric(df_plot_long$Cohens_d)
df_plot_long$SD<- as.numeric(df_plot_long$SD)

# Order approaches and add blank spaces for nicer plotting
approaches_ordered <- c("panic.trad\n(linear)", "panic.adv\n(ridge)", "panic.adv\n(ridgewhp)","panic.adv\n(rf)",  "gap1",              "protect.trad\n(linear)", "protect.adv\n(ridge)", "protect.adv\n(ridgewhp)","protect.adv\n(rf)", "gap2",  "aprotectsympscore.adv\n(ridge)", "aprotectfuncscore.adv\n(ridge)")

approaches_ordered_red <- c("panic.trad\n(linear)", "panic.adv\n(ridge)", "panic.adv\n(ridgewhp)","panic.adv\n(rf)",  "gap1",              "protect.trad\n(linear)", "protect.adv\n(ridge)", "protect.adv\n(ridgewhp)","protect.adv\n(rf)")

approaches_labels <- c("trad. (linear)", " adv. (ridge)", " adv. (ridge w. hp)"," adv. (rf)",  "",              "trad. (linear)", " adv. (ridge)", " adv. (ridge w. hp)"," adv. (rf)", "",  "adv.\n(symptom score, ridge)", "adv.\n(functioning score, ridge)")

approaches_labels_red <- c("trad. (linear)", " adv. (ridge)", " adv. (ridge w. hp)"," adv. (rf)",  "",              "trad. (linear)", " adv. (ridge)", " adv. (ridge w. hp)"," adv. (rf)")

# approaches_labels_red <- c("traditional (linear)", " advanced (ridge)", " advanced (ridge w. hp)"," advanced (random forest)",  "",              "traditional (linear)", " advanced (ridge)", " advanced (ridge w. hp)"," advanced (random forest)")


df_plot_long$approach <- factor(df_plot_long$approach,levels = approaches_ordered)

# Add type of approach for colours
df_plot_long$class <- ifelse(grepl("trad", df_plot_long$approach), "traditional", 
                        ifelse(grepl("adv", df_plot_long$approach), "advanced", NA))
df_plot_long$class <- factor(df_plot_long$class, levels = c("traditional","advanced"))

# Remove last classes
df_plot_long <- df_plot_long[!df_plot_long$approach %in% c("aprotectsympscore.adv\n(ridge)", "aprotectfuncscore.adv\n(ridge)"), ]
df_plot_long$approach <- factor(df_plot_long$approach, approaches_ordered_red)
```

## Setting for ggplot
```{r}
size_text = 24
theme_set(theme_bw() + 
            theme(panel.grid = element_blank(),
                  panel.grid.major.y = element_line(size = 0.2, color = "gray")    # Add only horizontal lines
)+ theme(text = element_text(size =size_text),
                             axis.title = element_text(size =size_text+4), 
                              plot.title = element_text(size = size_text),
                             axis.text = element_text(size = size_text),
                             strip.text = element_text(size = size_text),
                             legend.text =  element_text(size =size_text-1)))

# Set colors
# Get 5 colors from the orange spectrum
colrs_red_green = c("red3","chartreuse3")
```

## Plot 
```{r, fig.height = 8}
plot <- ggplot(data=df_plot_long, aes(x=approach, y=Cohens_d, fill = class)) +
  geom_bar(stat="identity") +
  geom_errorbar(aes(ymin=Cohens_d-SD, ymax=Cohens_d+SD), width=.2,
                 position=position_dodge(.9))+
  theme(axis.text.x = element_text(angle = 45, hjust = 1))+
  scale_fill_manual(values = colrs_red_green)+
  scale_x_discrete(labels = approaches_labels_red, drop = FALSE) +
  theme(legend.position="none") +
  xlab("")+
  ylab("Cohen´s d")

plot
```

## Save plot 
```{r}
ggsave("ML_Cohens_d.svg",plot, width = 13, height = 7)
```