---
title: "PAI_Mini_Review"
author: "Charlotte Meinke"
date: '2023-11-15'
output: html_document
---

# General settings
## Load packages
```{r}
library(readxl)
library(tidyr)
library(ggplot2)
library(dplyr)
library(flextable)
library(officer)
library(readxl)
```

## Settings ggplot
```{r}
# Save all plots in the following path:
path_save_plots = "Y:\\PsyThera\\Projekte_Meinke\\PAI_Personalized_Advantage_Index\\Schreiben\\Plots"

size_text = 8 # for psychological medicine
family = "Arial"
theme_set(theme_classic() + theme(text = element_text(size = size_text, family = family),
                             axis.title = element_text(size = size_text, family = family), 
                             plot.title = element_text(size = size_text, family = family),
                             axis.text = element_text(size = size_text, family = family),
                             strip.text =  element_text(size = size_text, family = family),
                             legend.text =  element_text(size = size_text, family = family)))
```

## Settings flextable
```{r, echo = FALSE}
# Flextable
set_flextable_defaults(font.family = "Arial",
font.size = 8,
padding.bottom = 3,
padding.top = 3,
padding.left = 0.5,
paddings.right = 0.5,
#theme_fun = "theme_apa",
theme_fun = NULL,
text.align = "center",
line_spacing = 1.5)

# Word documents to save flextables
margins <- page_mar(
  bottom = 0.5,
  top = 0.5,
  right = 0.5,
  left = 0.5,
  header = 0.5,
  footer = 0.5,
  gutter = 0.5
)

format_table_wide <- prop_section(
  page_size = page_size(orient = "landscape"),
  page_margins = margins)
```

# Table 1: Study characteristics table

## Import data

```{r import excel, results = 'hide', warning = FALSE}
data_study_char <-
  readxl::read_excel(path = "Extraction_PAI_studies_final_CM_KH_CM.xlsx", sheet = "study_characteristics_table",
                     na = "NA",col_types = "text")
```

## Preprocess data

```{r summary table with study characteristics, warning = FALSE}
# # Extract column names from first row
# colnames(data_study_char) <- data_study_char[1,]
# data_study_char <- data_study_char[-1,]

# Delete empty rows
data_study_char <- data_study_char[!(is.na(data_study_char$`First Author, Year`)),]

# Delete "et al."
data_study_char$`First Author, Year` <- gsub(" et al.", "",data_study_char$`First Author, Year`)

# Exclude column "Most relevant predictors"
data_study_char$`Most relevant predictors` <- NULL

# Exclude column Comments
data_study_char$`Comments` <- NULL

data_study_char
```

## Create flextable with study characteristics

```{r flextable settings, echo=FALSE, warning = FALSE}

# Create flextable
ft_sumtable <- flextable(data_study_char)

# Change header labels
ft_sumtable <- set_header_labels(ft_sumtable, 
                                 "post-treatment severity measure" = "outcome measure",
                                 "type of feature selection approach" = "Feature selection",
                                 "type of outcome prediction approach" = "Prediction",
                                 "CV scheme" = "CV",
                                 "Cohen's d for optimal vs non-optimal" = "Cohen's d")

ft_sumtable
```

## Prepare and save table in word

```{r Add title, table description, footer, echo=FALSE, warning = FALSE}
# Set table properties
ft_sumtable <- set_table_properties(ft_sumtable, width = 1, layout = "autofit")

# Header in bold
ft_sumtable <- bold(ft_sumtable, bold = TRUE, part = "header")

# Export flextable
save_as_docx(
  ft_sumtable,
  path = file.path(path_save_plots,"Table_1_study_characteristics.docx"), pr_section = format_table_wide)
```

# Fig 2a: Visualize Cohen´s d for the whole sample and 50% with the highest PAI

## Import data
We use a slightly different table than the one used for the study characterics table as we want to have each subanalysis in one row.

```{r}
data_study_char <-
  readxl::read_excel(path = "Extraction_PAI_studies_final_CM_KH_CM.xlsx", sheet = "subanalyses_cohens_d", na = "NA")

start_index <- which(names(data_study_char) == "Mean absolute PAI")

# Convert columns to numeric, starting from the specified column
data_study_char[, start_index:ncol(data_study_char)] <- lapply(data_study_char[, start_index:ncol(data_study_char)], as.numeric)
```

## Subset data
```{r}
subsetted_data <- data_study_char[, c("First Author, Year", "Cohen's d for optimal vs non-optimal", "Subsample: Cohen's d for optimal vs non-optimal")]
colnames(subsetted_data) <- c("Authors", "Entire sample", "Subsample")

# Order after NA (CAVE: excel-file needs to be ordered alphabetically, to get the desired order)
subsetted_data <- subsetted_data[order(is.na(subsetted_data$`Subsample`)), ]
```

## Prepare data for plotting
```{r}
# Convert data to long format and omit missing values (NAs)
long_data <- subsetted_data %>%
  tidyr::gather(key = "Sample_Type", value = "Cohen_D", -Authors) %>%
  na.omit()

# Convert Cohen_D to numeric
long_data$Cohen_D <- as.numeric(long_data$Cohen_D)
mean_Cohen_all = mean(subsetted_data$`Entire sample`,na.rm=TRUE)
mean_Cohen_sub = mean(subsetted_data$`Subsample`,na.rm=TRUE)

# Improve study names
long_data$Authors <- gsub(" subanalysis", "",long_data$Authors)
long_data$Authors <- gsub(" et al.", "",long_data$Authors)

# Here, the variable needs only to be reversed.
long_data$Authors <- factor(long_data$Authors, levels = rev(unique(long_data$Authors)))
```

## Plot data
```{r}
# Dot plot
#colour_palette = c("#0CEBE1", "#3C8581")
colour_palette = c("#666666","#CCCCCC")
lev = levels(long_data$Authors)

#cohen_1 = as.vector(long_data[long_data$Authors == lev[18],c("Cohen_D")])$Cohen_D[1]
#cohen_2 = as.vector(long_data[long_data$Authors == lev[18],c("Cohen_D")])$Cohen_D[2]

plot_2a <- ggplot(long_data, aes(x = Authors, y = Cohen_D, color = Sample_Type)) +
  geom_point(position = position_dodge(width = 0), size = 3) +
  geom_hline(yintercept = mean_Cohen_all, color = colour_palette[1]) + 
  geom_segment(aes(x = Inf, xend = "Webb, 2019", y = mean_Cohen_sub, yend = mean_Cohen_sub), color = colour_palette[2]) +
  # geom_segment(aes(x = lev[18], xend = lev[18], 
  #                  y = as.vector(long_data[long_data$Authors == lev[18],c("Cohen_D")])$Cohen_D[1], 
  #                  yend = as.vector(long_data[long_data$Authors == lev[18],c("Cohen_D")])$Cohen_D[2]), 
  #              color = "grey") +
  labs(
    x = "",
    y = "Cohen's d",
    color = "Sample Type"
  ) +
  scale_y_continuous(
    breaks = c(0,0.1,0.2,0.3,0.4,0.5,0.6,0.7,0.8),
    limits = c(0, 0.8)
  ) + 
  scale_color_manual(values = colour_palette)+ 
  theme_bw() + 
  theme(text = element_text(size = size_text, family = family),
                             axis.title = element_text(size = size_text, family = family), 
                             plot.title = element_text(size = size_text, family = family),
                             axis.text = element_text(size = size_text, family = family),
                             strip.text =  element_text(size = size_text, family = family),
                             legend.text =  element_text(size = size_text, family = family)) +
  theme(
    panel.grid = element_blank(),          # Remove all grid lines
    panel.grid.major.y = element_line(size = 0.2, color = "gray")    # Add only horizontal lines
  )+
  coord_flip() +
  theme(legend.position = "bottom", legend.title = element_blank())
  #theme(axis.text.x = element_text(angle = 45, hjust = 1))

# # Violin Plot 
# ggplot(long_data, aes(x = Sample_Type, y = Cohen_D, fill = Sample_Type)) +
#   geom_violin() +
#   labs(
#     title = "Distribution of Cohen's D by Sample Type",
#     x = "Sample Type",
#     y = "Cohen's D",
#     fill = "Sample Type"
#   ) +
#   theme_minimal()

plot_2a

```

# PROBAST-Rating
## Import data
```{r}
data_PROBAST <- readxl::read_excel(path = "PAI_PROBAST_assessment_KH_CM.xlsx", sheet = "Main")
colnames(data_PROBAST) <- data_PROBAST[1,]
data_PROBAST <- data_PROBAST[2:nrow(data_PROBAST),]
```

## Table S1
```{r}
# Prepare dataset
# Remove columns that contain "comments"
cols_no_comments <- colnames(data_PROBAST)[!grepl("comments",colnames(data_PROBAST))]
data_PROBAST_table <- data_PROBAST[,cols_no_comments]
# Remove "et al."
data_PROBAST_table$Study <- gsub(" et al.", "",data_PROBAST_table$Study)

# Transpose table
PROBAST_transposed <- t(data_PROBAST_table)
PROBAST_transposed <- as.data.frame(PROBAST_transposed)
colnames(PROBAST_transposed) <- PROBAST_transposed[1, ]
PROBAST_transposed <- PROBAST_transposed[-1, ]
# Use rownames as first column (as flextable ignores rownames)
PROBAST_transposed <- PROBAST_transposed %>%
  tibble::rownames_to_column(var = "Question")

# Combine results for subanalyses
colnames(PROBAST_transposed)[colnames(PROBAST_transposed)=="Hoeboer, 2021; subanalysis CAPS-5"] <- "Hoeboer, 2021; CAPS-5, PCL-5"
PROBAST_transposed$`Hoeboer, 2021; subanalysis PCL-5` <- NULL
colnames(PROBAST_transposed)[colnames(PROBAST_transposed)=="van Bronswijk, 2021; subanalysis FreqMesh"] <- "van Bronswijk, 2021; FreqMesh, STEPd"
PROBAST_transposed$`van Bronswijk, 2021; subanalysis STEPd` <- NULL
colnames(PROBAST_transposed)[colnames(PROBAST_transposed)=="van Bronswijk, 2021; subanalysis STEPd to FreqMesh"] <- "van Bronswijk, 2021; cross-trial"
PROBAST_transposed$`van Bronswijk, 2021; subanalysis FreqMesh to STEPd` <- NULL

# Generate flextable
ft <- flextable(PROBAST_transposed)

ft <- set_table_properties(ft, width = 1, layout = "autofit")
ft <- autofit(ft)
ft <- hline(ft, part = "body", i = c(3, 7, 14, 26), 
                      border = fp_border(color = "black", width = 1))
ft <- fontsize(ft, size = 7, part = "all")

flextable::save_as_docx(ft, path = file.path(path_save_plots,"PROBAST_table.docx"), pr_section = format_table_wide)
```

## Plot S1
```{r Plot_PROBAST}
# Final rating columns
cols_final_rating <- c("Final rating domain 1 (risk of bias)", 
                       "Final rating domain 2 (risk of bias)", 
                       "Final rating domain 3 (risk of bias)", 
                       "Final rating domain 4 (risk of bias)", 
                       "Final rating (risk of bias)")

data_PROBAST_plot <- data_PROBAST[c("Study",cols_final_rating)]

# Bring data into long-format
data_PROBAST_plot_long <- reshape(data = data_PROBAST_plot,idvar = "Study", new.row.names = 1:20000,varying = cols_final_rating, v.name = "ROB", times = cols_final_rating, direction = "long")
colnames(data_PROBAST_plot_long)[colnames(data_PROBAST_plot_long)=="time"] <- 'Rating_domain'

# Order and rename factor rating_domain
PROBAST_domains_eng <- c("ROB Sample","ROB Predictors", "ROB Outcome", "ROB Analysis", "ROB Total")
data_PROBAST_plot_long$Rating_domain <- factor(data_PROBAST_plot_long$Rating_domain, 
                                               levels = cols_final_rating, 
                                               labels = PROBAST_domains_eng)

scaleFUN <- function(x) x*100

# Prepare final rating domains labels to make one label bold
breaks <- levels(data_PROBAST_plot_long$Rating_domain)
labels <- as.expression(breaks)
labels[[5]] <- bquote(bold(.(labels[[5]])))

# Plot data
legend_PROBAST_eng <- c("high", "low")
plot_PROBAST <- ggplot(data = data_PROBAST_plot_long,aes(x = `Rating_domain`, fill = `ROB`))+ geom_bar(aes (alpha = Rating_domain == "ROB Total", y = ..count../tapply(..count.., ..x.. ,sum)[..x..]))+
  coord_flip()+
  scale_fill_manual(values = c("red3","chartreuse3","grey"), name = "Risk of bias (ROB)", labels = legend_PROBAST_eng <- c("high", "low")) +
  scale_y_continuous(labels=scaleFUN)+
  scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.6), guide = "none")+
  scale_x_discrete(label = labels, breaks = breaks)+
   theme(legend.position = "top")+
  xlab("")+
  ylab("Relative Proportion in %")

# Save plot
plot_PROBAST
ggsave(file.path(path_save_plots,"figure_S1_PROBAST.svg"),plot_PROBAST, height = 4.5, width = 9)
```

## Figure 2b: Plot ROB vs. Cohen´s d

### Prepare data
```{r}
# Bring data into long-format
cols_imp <- c("4.8.1 What is the extent of ROB introduced by the cross-validation procedure?","4.8.2 What is the extent of ROB introduced by (not) integrating preprocessing steps into the CV?")
data_PROBAST_48 <- data_PROBAST[c("Study",cols_imp)]

# Add Cohen´s d from study characteristics data (merge by first column)
data_study_char_imp <- data_study_char[,c("First Author, Year","Cohen's d for optimal vs non-optimal")]
colnames(data_study_char_imp)[colnames(data_study_char_imp) == "First Author, Year"] <- "Study"
data_Cohen_PROBAST <- merge(data_PROBAST_48, data_study_char_imp, by = "Study")

# Change column names
colnames(data_Cohen_PROBAST) <- c("Study","4.8.1_cv","4.8.2_steps","cohens_d")

data_Cohen_PROBAST
```

### Prepare data for plotting
```{r echo=TRUE}
# Bring data into long-format
data_Cohen_PROBAST_long <- reshape(data = data_Cohen_PROBAST,idvar = "Study", new.row.names = 1:20000,varying = c("4.8.1_cv","4.8.2_steps"), v.name = "ROB", times = c("4.8.1_cv","4.8.2_steps"), direction = "long")
colnames(data_Cohen_PROBAST_long)[colnames(data_Cohen_PROBAST_long)=="time"] <- 'Question'

data_Cohen_PROBAST_long$ROB <- gsub(" ROB","",data_Cohen_PROBAST_long$ROB)
data_Cohen_PROBAST_long$ROB <- factor(data_Cohen_PROBAST_long$ROB, levels = c("low","medium", "high"))

data_Cohen_PROBAST_long$Question <- dplyr::recode(data_Cohen_PROBAST_long$Question, "4.8.1_cv" = "Cross-validation", "4.8.2_steps" = "Data leakage")

# Add mean column to depict mean in legend (workaround)
data_Cohen_PROBAST_long$mean <- rep("mean",nrow(data_Cohen_PROBAST_long))
```

### Plot
```{r}
data_Cohen_PROBAST_long$cohens_d <- as.numeric(data_Cohen_PROBAST_long$cohens_d )
plot_2b <- ggplot(data_Cohen_PROBAST_long, aes(x = ROB, y = cohens_d)) +
  geom_jitter(size = 2, width = 0.1)+
  stat_summary(fun = "mean", fun.min = "mean", fun.max= "mean", size= 0.3, geom = "crossbar")+
  ylab("Cohen´s d")+
  xlab("Risk of bias")+
  #geom_segment(aes(xend = ROB, yend = cohens_d, color = "mean"), size = 3)+
  geom_point(aes(shape = mean),size = 10, alpha = 0)+
  scale_shape_manual(values = c("-"))+
  guides(shape=guide_legend(override.aes = list(alpha = 1)))+
  theme(legend.position = "bottom", legend.title = element_blank())+
  facet_wrap(~Question)

plot_2b + theme(
  #strip.background = element_blank(),
  strip.placement = "above" + theme_bw()
)
```

# Combine Plot 2a and 2b
```{r}
library(patchwork)

plot_2 <- plot_2a + plot_2b + plot_annotation(tag_levels = "A")

# Savings for psychological medicine: dpi = 800
# color: tif; grey + linework: eps
ggsave(file.path(path_save_plots,"Figure_2_Cohens_d.svg"), plot_2, units = "cm", width = 20.5, height = 12, dpi = 800)
ggsave(file.path(path_save_plots,"Figure_2_Cohens_d.eps"), plot_2, units = "cm", width = 20.5, height = 12, dpi = 800)
```
